---
title: "CRISPRi PHENOMICS DATA ANALYSIS"
author: "Vaskar Mukherjee"
date: "2/3/2021"
output: 
  github_document:
    toc: true
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# IMPORT SCAN-O-MATIC RAW DATA

The phenotypic data generated in scan-o-matic screening in .csv format. 
We extract both the absolute and the normalized phenotypes.

The CRISPRi strains in the library were arrayed in 24 plates in 384 format. Each CRISPRi plate was subjected to two different condition (Basal and 150 mM of Acetic acid). Therefore, for each plate four different files are generated. All files generated in a single independent experimental round are stored in a single folder. 

* **SOM_SCR_R001** : Raw data for round1

* **SOM_SCR_R002** : Raw data for round2

**ABSOLUTE DATA**

The **Absolute** dataset gives the extracted phenotypes without any spatial normalization  

**NORMALIZED DATA**

The **Normalized** dataset is generated after removal of any spatial bias. This is in log2 scale and referred as Log Strain Coefficient (LSC) values

**FILE NAMING**

Each file is named with the plate identifier in such a way so that it can be easily called programmatically  

Eg. Plate 1 absolute data in basal (Ctrl) condition have the following string  
*Ctrl1.phenotypes.Absolute*  
AND  
Plate 1 Normalized data in acetic acid (aa) stress have the string  
*aa1.phenotypes.Normalized*

## PURPOSE 1
At the end of this data import session, a single data.frame will be generated with the data of 24 plates. The whole dataset will be labeled with the strains attributes using the metadata key file (provided in the COMPILED_DATA folder). The data import below is shown for only Round2 dataset. Round1 can be generated modifying the folder location

**METADATA KEY FILE** : library_keyfile1536.csv

## IMPORTING THE METADATA FILE  

```{r}
Metadata_CRISPRi <- read.csv("COMPILED_DATA/library_keyfile1536.csv", na.strings = "#N/A", stringsAsFactors = FALSE)
str(Metadata_CRISPRi)
```

## GENERATE BASAL **ABSOLUTE** DATASET

```{r}
m <- vector(mode = "character", length = 0)
file.names<-vector(mode = "character", length = 0)
temp_df<-data.frame()
data_Ctrl_Abs <- data.frame()
for(i in 1:24){
  m <- paste0("Ctrl", i, ".phenotypes.Absolute") 
  file.names[i] <- dir("RAW_DATA/SOM_SCR_R002/", pattern = m, full.names = TRUE)
  temp_df <- read.csv(file.names[i], na.strings = "NoGrowth")
  data_Ctrl_Abs <- rbind(data_Ctrl_Abs, temp_df)
}
str(data_Ctrl_Abs)
```

Several phenotypes are extracted. However, the most useful for this study will be, 

* Column No: 14 i.e. **Phenotypes.ExperimentGrowthYield**
* Column No: 15 i.e. **Phenotypes.GenerationTime**

Extract only this two column in the final data.frame  
Rename the column names to prevent any ambiguity  

```{r}
data_Ctrl_Abs_Trim <- data_Ctrl_Abs[, 14:15]
colnames(data_Ctrl_Abs_Trim) <- c("CTRL_Y_ABS", "CTRL_GT_ABS")
str(data_Ctrl_Abs_Trim)
```

## GENERATE ACETIC ACID **ABSOLUTE** DATASET

Following the same strategy as above

```{r}
m <- vector(mode = "character", length = 0)
file.names<-vector(mode = "character", length = 0)
temp_df<-data.frame()
data_AA_Abs <- data.frame()
for(i in 1:24){
  m <- paste0("aa", i, ".phenotypes.Absolute") 
  file.names[i] <- dir("RAW_DATA/SOM_SCR_R002/", pattern = m, full.names = TRUE)
  temp_df <- read.csv(file.names[i], na.strings = "NoGrowth")
  data_AA_Abs <- rbind(data_AA_Abs, temp_df)
}
data_AA_Abs_Trim <- data_AA_Abs[, 14:15]
colnames(data_AA_Abs_Trim) <- c("AA_Y_ABS", "AA_GT_ABS")
```

## GENERATE BASAL **NORMALIZED** DATASET

```{r}
m <- vector(mode = "character", length = 0)
file.names<-vector(mode = "character", length = 0)
temp_df<-data.frame()
data_Ctrl_Norm <- data.frame()
for(i in 1:24){
  m <- paste0("Ctrl", i, ".phenotypes.Normalized") 
  file.names[i] <- dir("RAW_DATA/SOM_SCR_R002/", pattern = m, full.names = TRUE)
  temp_df <- read.csv(file.names[i], na.strings = "NoGrowth")
  data_Ctrl_Norm <- rbind(data_Ctrl_Norm, temp_df)
}
str(data_Ctrl_Norm)
```

The most useful for this study will be,

* Column No: 4 i.e. Phenotypes.ExperimentGrowthYield 
* Column No: 5 i.e. Phenotypes.GenerationTime

Extract only this two column  

```{r}
data_Ctrl_Norm_Trim <- data_Ctrl_Norm[, 4:5]
colnames(data_Ctrl_Norm_Trim) <- c("CTRL_Y_NORM", "CTRL_GT_NORM")
```

## GENERATE ACETIC ACID **NORMALIZED** DATASET

Same as above 

```{r}
m <- vector(mode = "character", length = 0)
file.names<-vector(mode = "character", length = 0)
temp_df<-data.frame()
data_AA_Norm <- data.frame()
for(i in 1:24){
  m <- paste0("aa", i, ".phenotypes.Normalized") 
  file.names[i] <- dir("RAW_DATA/SOM_SCR_R002/", pattern = m, full.names = TRUE)
  temp_df <- read.csv(file.names[i], na.strings = "NoGrowth")
  data_AA_Norm <- rbind(data_AA_Norm, temp_df)
}
data_AA_Norm_Trim <- data_AA_Norm[, 4:5]
colnames(data_AA_Norm_Trim) <- c("AA_Y_NORM", "AA_GT_NORM")
```

## COMBINE THE DATASETS TO OBTAIN FINAL DATAFRAME

Trimmed datasets are combined to obtain the final data.frame.
The combined data frame is labeled as data from ROUND2

```{r}
R <- rep("2nd_round", 36864)
Round_ID <- data.frame(R, stringsAsFactors = FALSE)
whole_data_R2 <- cbind(Metadata_CRISPRi, 
                       Round_ID, 
                       data_Ctrl_Abs_Trim, 
                       data_AA_Abs_Trim, 
                       data_Ctrl_Norm_Trim, 
                       data_AA_Norm_Trim)
colnames(whole_data_R2)[12] <- "Round_ID"
str(whole_data_R2)
```

## IMPORT RESULTS FROM ROUND1

The results from Round1 is already compiled to a .csv file in COMPILED_DATA folder
**Results 1st Round** : 20190903_CRISPRi_Screen_aa_1st_round.csv

Import the dataset and label as data from ROUND1

```{r}
First_round <- read.csv("COMPILED_DATA/20190903_CRISPRi_Screen_aa_1st_round.csv", 
                        na.strings = c("#N/A", "NoGrowth"), 
                        stringsAsFactors = FALSE)
R <- rep("1st_round", 36864)
Round_ID <- data.frame(R, stringsAsFactors = FALSE)
whole_data_R1 <- cbind(Metadata_CRISPRi, Round_ID, First_round[, 12:19])
colnames(whole_data_R1)[12] <- "Round_ID"
str(whole_data_R1)
```

## COMBINE THE DATASETS of ROUND 1 AND 2

```{r}
whole_data_CRISPRi_aa <- rbind(whole_data_R1, whole_data_R2)
```

# SCAN-O-MATIC PHENOMICS ANALYSIS 

In this study most of the downstream analysis was performed using the phenotype Generation_time(GT)

## PURPOSE 2

In this session, downstream data processing and statistical analysis of SCAN-O-MATIC raw output will be performed

## ESTIMATE THE LOG PHENOTYPIC INDEX (LPI) VALUES

LPI of strain is the difference of its normalized Generation_Time(GT) / Yield(Y) (LSC, see [IMPORT SCAN-O-MATIC RAW DATA]) on acetic acid stress plate to the basal condition. It gives a **RELATIVE** estimate of how a strain performed under acetic acid stress relative to the basal condition.   

The **RELATIVE GENERATION TIME** i.e. LPI_GT = LSC_GT_Acetic_Acid - LSC_GT_Basal 

```{r}
whole_data_CRISPRi_aa[, 21] <- whole_data_CRISPRi_aa[, 19]-whole_data_CRISPRi_aa[, 17]
whole_data_CRISPRi_aa[, 22] <- whole_data_CRISPRi_aa[, 20]-whole_data_CRISPRi_aa[, 18]
colnames(whole_data_CRISPRi_aa)[21] <- "LPI_Y"
colnames(whole_data_CRISPRi_aa)[22] <- "LPI_GT"
```

## PERFORM PLATE-WISE BATCH CORRECTION

Plate-wise batch correction was conducted by subtracting the median of LSC GT values of all the individual colonies on a plate from the individual LSC GT values of the colonies growing on that plate. 

i.e. if strainX is growing in Basal condition on plate Z, the corrected LSC_GT value for strainX in the Basal condition is the following;

* LSC_GT_Basal_Corrected~strainX~ = (LSC_GT_Basal~strainX~) - Median(LSC_GT Basal~PlateZ~)

```{r}
plate_ID <- as.character(unique(whole_data_CRISPRi_aa$SOURCEPLATEID))
whole_data_CRISPRi_aa_corrected <- whole_data_CRISPRi_aa
med_LogLSCctrl_RND1_GT <- vector(mode = "integer", length = 0)
med_LogLSCaa_RND1_GT <- vector(mode = "integer", length = 0)
med_LogLSCctrl_RND2_GT <- vector(mode = "integer", length = 0)
med_LogLSCaa_RND2_GT <- vector(mode = "integer", length = 0)
med_LogLSCctrl_RND1_Y <- vector(mode = "integer", length = 0)
med_LogLSCaa_RND1_Y <- vector(mode = "integer", length = 0)
med_LogLSCctrl_RND2_Y <- vector(mode = "integer", length = 0)
med_LogLSCaa_RND2_Y <- vector(mode = "integer", length = 0)

for(i in 1:24){
med_LogLSCctrl_RND1_GT[i] <- median(whole_data_CRISPRi_aa_corrected$CTRL_GT_NORM[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i]
                                                                                       & !is.na(whole_data_CRISPRi_aa_corrected$CTRL_GT_NORM) 
                                                                                         & whole_data_CRISPRi_aa_corrected$Round_ID=="1st_round")])

med_LogLSCaa_RND1_GT[i] <- median(whole_data_CRISPRi_aa_corrected$AA_GT_NORM[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i]
                                                                                     & !is.na(whole_data_CRISPRi_aa_corrected$AA_GT_NORM) 
                                                                                     & whole_data_CRISPRi_aa_corrected$Round_ID=="1st_round")])

med_LogLSCctrl_RND2_GT[i] <- median(whole_data_CRISPRi_aa_corrected$CTRL_GT_NORM[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i]
                                                                                         & !is.na(whole_data_CRISPRi_aa_corrected$CTRL_GT_NORM) 
                                                                                         & whole_data_CRISPRi_aa_corrected$Round_ID=="2nd_round")])

med_LogLSCaa_RND2_GT[i] <- median(whole_data_CRISPRi_aa_corrected$AA_GT_NORM[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i]
                                                                                     & !is.na(whole_data_CRISPRi_aa_corrected$AA_GT_NORM) 
                                                                                     & whole_data_CRISPRi_aa_corrected$Round_ID=="2nd_round")])

med_LogLSCctrl_RND1_Y[i] <- median(whole_data_CRISPRi_aa_corrected$CTRL_Y_NORM[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i]
                                                                                       & !is.na(whole_data_CRISPRi_aa_corrected$CTRL_Y_NORM) 
                                                                                       & whole_data_CRISPRi_aa_corrected$Round_ID=="1st_round")])

med_LogLSCaa_RND1_Y[i] <- median(whole_data_CRISPRi_aa_corrected$AA_Y_NORM[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i]
                                                                                   & !is.na(whole_data_CRISPRi_aa_corrected$AA_Y_NORM) 
                                                                                   & whole_data_CRISPRi_aa_corrected$Round_ID=="1st_round")])

med_LogLSCctrl_RND2_Y[i] <- median(whole_data_CRISPRi_aa_corrected$CTRL_Y_NORM[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i]
                                                                                       & !is.na(whole_data_CRISPRi_aa_corrected$CTRL_Y_NORM) 
                                                                                       & whole_data_CRISPRi_aa_corrected$Round_ID=="2nd_round")])

med_LogLSCaa_RND2_Y[i] <- median(whole_data_CRISPRi_aa_corrected$AA_Y_NORM[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i]
                                                                                   & !is.na(whole_data_CRISPRi_aa_corrected$AA_Y_NORM) 
                                                                                   & whole_data_CRISPRi_aa_corrected$Round_ID=="2nd_round")])
  
whole_data_CRISPRi_aa_corrected[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i] &
                                          whole_data_CRISPRi_aa_corrected$Round_ID=="1st_round") , 23] <- whole_data_CRISPRi_aa_corrected[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i] &
                                                                                                                                                  whole_data_CRISPRi_aa_corrected$Round_ID=="1st_round"), 17] - med_LogLSCctrl_RND1_Y[i]
  whole_data_CRISPRi_aa_corrected[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i] &
                                          whole_data_CRISPRi_aa_corrected$Round_ID=="1st_round") , 24] <- whole_data_CRISPRi_aa_corrected[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i] &
                                                                                                                                                  whole_data_CRISPRi_aa_corrected$Round_ID=="1st_round"), 18] - med_LogLSCctrl_RND1_GT[i]
  whole_data_CRISPRi_aa_corrected[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i] &
                                          whole_data_CRISPRi_aa_corrected$Round_ID=="2nd_round") , 23] <- whole_data_CRISPRi_aa_corrected[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i] &
                                                                                                                                                  whole_data_CRISPRi_aa_corrected$Round_ID=="2nd_round"), 17] - med_LogLSCctrl_RND2_Y[i]
  whole_data_CRISPRi_aa_corrected[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i] &
                                          whole_data_CRISPRi_aa_corrected$Round_ID=="2nd_round") , 24] <- whole_data_CRISPRi_aa_corrected[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i] &
                                                                                                                                                  whole_data_CRISPRi_aa_corrected$Round_ID=="2nd_round"), 18] - med_LogLSCctrl_RND2_GT[i]
  whole_data_CRISPRi_aa_corrected[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i] &
                                          whole_data_CRISPRi_aa_corrected$Round_ID=="1st_round") , 25] <- whole_data_CRISPRi_aa_corrected[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i] &
                                                                                                                                                  whole_data_CRISPRi_aa_corrected$Round_ID=="1st_round"), 19] - med_LogLSCaa_RND1_Y[i]
  whole_data_CRISPRi_aa_corrected[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i] &
                                          whole_data_CRISPRi_aa_corrected$Round_ID=="1st_round") , 26] <- whole_data_CRISPRi_aa_corrected[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i] &
                                                                                                                                                  whole_data_CRISPRi_aa_corrected$Round_ID=="1st_round"), 20] - med_LogLSCaa_RND1_GT[i]
  whole_data_CRISPRi_aa_corrected[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i] &
                                          whole_data_CRISPRi_aa_corrected$Round_ID=="2nd_round") , 25] <- whole_data_CRISPRi_aa_corrected[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i] &
                                                                                                                                                  whole_data_CRISPRi_aa_corrected$Round_ID=="2nd_round"), 19] - med_LogLSCaa_RND2_Y[i]
  whole_data_CRISPRi_aa_corrected[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i] &
                                          whole_data_CRISPRi_aa_corrected$Round_ID=="2nd_round") , 26] <- whole_data_CRISPRi_aa_corrected[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i] &
                                                                                                                                                  whole_data_CRISPRi_aa_corrected$Round_ID=="2nd_round"), 20] - med_LogLSCaa_RND2_GT[i]
}
```

## ESTIMATE THE BATCH CORRECTED LOG PHENOTYPIC INDEX (LPI) VALUES

Estimate the corrected LPI values (see [ESTIMATE THE LOG PHENOTYPIC INDEX (LPI) VALUES]) based on the corrected LSC values

i.e. LPI_GT~corrected~ = LSC_GT_Acetic_Acid~corrected~ - LSC_GT_Basal~corrected~

**Estimate the corrected LPI_Y**
```{r}
whole_data_CRISPRi_aa_corrected[, 27] <- whole_data_CRISPRi_aa_corrected[, 25] - whole_data_CRISPRi_aa_corrected[, 23]
```

**Estimate the corrected LPI_GT**
```{r}
whole_data_CRISPRi_aa_corrected[, 28] <- whole_data_CRISPRi_aa_corrected[, 26] - whole_data_CRISPRi_aa_corrected[, 24] 
```

## SETTING THE NAMES OF THE NEW COLUMNS

```{r}
colnm <- colnames(whole_data_CRISPRi_aa)[17:22]
colnm <- paste0(colnm, "_CR")
colnames(whole_data_CRISPRi_aa_corrected)[23:28] <- colnm
str(whole_data_CRISPRi_aa_corrected)
```

## EXTRACT ONLY THE BATCH CORRECTED COLUMNS

```{r}
whole_data_CRISPRi_aa_2 <- whole_data_CRISPRi_aa_corrected[, c(1:16, 23:28)]
colnames(whole_data_CRISPRi_aa_2)[17:22] <- colnames(whole_data_CRISPRi_aa)[17:22]
str(whole_data_CRISPRi_aa_2)
```

## CONSTRUCT A NEW DATA STRUCTURE

Construct a new data structure where data from each strain (have a unique guide-RNA) is in a separate row and the replicates from first and second round are side by side. Add also the mean, median and standard deviation statistics for each phenotype  

### REMOVE ROWS WITH SPATIAL CONTROL STRAIN DATA

```{r}
Data_CRISPRi_aa <- subset(whole_data_CRISPRi_aa_2, whole_data_CRISPRi_aa_2$gRNA_name!="SP_Ctrl_CC23")
```

### CREATE A TABLE OF UNIQUE gRNA

```{r}
df_unique_sgRNA <- data.frame(table(Data_CRISPRi_aa$gRNA_name))
```

### ARRANGE THE DATA IN THE DESIRED FORMAT

```{r}
R1<-vector(mode = "integer", length = 0)
R2<-vector(mode = "integer", length = 0)
test2<-data.frame()
n<-nrow(df_unique_sgRNA)
for(i in 1:n){
  R1 <- which(Data_CRISPRi_aa$gRNA_name==df_unique_sgRNA$Var1[i] & Data_CRISPRi_aa$Round_ID=="1st_round")
  R2 <- which(Data_CRISPRi_aa$gRNA_name==df_unique_sgRNA$Var1[i] & Data_CRISPRi_aa$Round_ID=="2nd_round")
  test1 <- Data_CRISPRi_aa[c(R1, R2), ]
  test2[i, c(1:8)]<-test1[1, c(2:4, 6:7, 9:11)]
  test2[i, c(9:14)] <- test1$CTRL_GT_NORM
  test2[i, 15] <- mean(test1$CTRL_GT_NORM[1:3])
  test2[i, 16] <- mean(test1$CTRL_GT_NORM[4:6])
  test2[i, 17] <- sd(test1$CTRL_GT_NORM[1:3])
  test2[i, 18] <- sd(test1$CTRL_GT_NORM[4:6])
  test2[i, 19] <- mean(test1$CTRL_GT_NORM[1:6])
  test2[i, 20] <- median(test1$CTRL_GT_NORM[1:6])
  test2[i, 21] <- sd(test1$CTRL_GT_NORM[1:6])
  test2[i, c(22:27)] <- test1$AA_GT_NORM
  test2[i, 28] <- mean(test1$AA_GT_NORM[1:3])
  test2[i, 29] <- mean(test1$AA_GT_NORM[4:6])
  test2[i, 30] <- sd(test1$AA_GT_NORM[1:3])
  test2[i, 31] <- sd(test1$AA_GT_NORM[4:6])
  test2[i, 32] <- mean(test1$AA_GT_NORM[1:6])
  test2[i, 33] <- median(test1$AA_GT_NORM[1:6])
  test2[i, 34] <- sd(test1$AA_GT_NORM[1:6])
  test2[i, c(35:40)] <- test1$LPI_GT
  test2[i, 41] <- mean(test1$LPI_GT[1:3])
  test2[i, 42] <- mean(test1$LPI_GT[4:6])
  test2[i, 43] <- sd(test1$LPI_GT[1:3])
  test2[i, 44] <- sd(test1$LPI_GT[4:6])
  test2[i, 45] <- mean(test1$LPI_GT[1:6])
  test2[i, 46] <- median(test1$LPI_GT[1:6])
  test2[i, 47] <- sd(test1$LPI_GT[1:6])
  test2[i, c(48:53)] <- test1$CTRL_Y_NORM
  test2[i, 54] <- mean(test1$CTRL_Y_NORM[1:3])
  test2[i, 55] <- mean(test1$CTRL_Y_NORM[4:6])
  test2[i, 56] <- sd(test1$CTRL_Y_NORM[1:3])
  test2[i, 57] <- sd(test1$CTRL_Y_NORM[4:6])
  test2[i, 58] <- mean(test1$CTRL_Y_NORM[1:6])
  test2[i, 59] <- median(test1$CTRL_Y_NORM[1:6])
  test2[i, 60] <- sd(test1$CTRL_Y_NORM[1:6])
  test2[i, c(61:66)] <- test1$AA_Y_NORM
  test2[i, 67] <- mean(test1$AA_Y_NORM[1:3])
  test2[i, 68] <- mean(test1$AA_Y_NORM[4:6])
  test2[i, 69] <- sd(test1$AA_Y_NORM[1:3])
  test2[i, 70] <- sd(test1$AA_Y_NORM[4:6])
  test2[i, 71] <- mean(test1$AA_Y_NORM[1:6])
  test2[i, 72] <- median(test1$AA_Y_NORM[1:6])
  test2[i, 73] <- sd(test1$AA_Y_NORM[1:6])
  test2[i, c(74:79)] <- test1$LPI_Y
  test2[i, 80] <- mean(test1$LPI_Y[1:3])
  test2[i, 81] <- mean(test1$LPI_Y[4:6])
  test2[i, 82] <- sd(test1$LPI_Y[1:3])
  test2[i, 83] <- sd(test1$LPI_Y[4:6])
  test2[i, 84] <- mean(test1$LPI_Y[1:6])
  test2[i, 85] <- median(test1$LPI_Y[1:6])
  test2[i, 86] <- sd(test1$LPI_Y[1:6])
}
```

### ASSIGN COLUMN NAMES 

Column names are already stored in a text times available in the **COMPILED_DATA** folder. Then store the data.frame under a new name.  
```{r}
column_names <- read.table("COMPILED_DATA/Column_names.txt", header = FALSE, sep = "\t", as.is = TRUE)
colnames(test2) <- column_names$V1
Analysis_CRISPRi_aa_Complete <- test2
str(Analysis_CRISPRi_aa_Complete)
```

## PERFORM STATISTICAL ANALYSIS

Multiple statistical method was applied to identify the best fit statistical model for this dataset. We start with the complete dataset and give it a new name to avoid distorting the original dataset.

```{r}
Analysis_Final <- Analysis_CRISPRi_aa_Complete
```


### METHOD 1

For **METHOD 1**, We hypothesized that the difference between the mean(µ) phenotypic performance of a specific CRISPRi strain (StrainX) in the two independent experimental rounds (n=2) to the mean phenotypic performance of all the CRISPRi strains that falls within the interquartile range (IQR) of the complete dataset would be zero, and any difference within the IQR to be just by chance. 

**Null Hypothesis** : µ(µ~LPI_GT_StrainX_Round1~, µ~LPI_GT_StrainX_Round2~)- µ(InterquartileRange_LPI_GT) = 0

#### RECALCULATION OF SOME PHENOTYPIC PARAMETERS

In this method, we estimate the Mean / standard deviation (SD) of the LPI GT of Round 1 and Round 2 separately for each strain. When one/two of the three replicates of a strain in a round returned missing value (i.e. NA), then the mean / SD of LPI GT for that round is calculated by taking average of the non NA replicates. Therefore, excluding the missing values the mean and SD statistics were recalculated. We implemented a if else decision tree for this  

* The mean and SD of Normalized generation time (**LSC GT mean**) at **Basal condition** re-calculation

```{r}
for(i in 1:nrow(Analysis_Final)){
  x1 <- as.numeric(Analysis_Final[i, 9:11][which(!is.na(Analysis_Final[i, 9:11]))])
  x2 <- as.numeric(Analysis_Final[i, 12:14][which(!is.na(Analysis_Final[i, 12:14]))])
  if(length(x1)==0){
    Analysis_Final$CTRL_GT_RND1_MEAN[i] <- NA
  } else{
    Analysis_Final$CTRL_GT_RND1_MEAN[i] <- as.numeric(mean(x1))
  }
  if(length(x2)==0){
    Analysis_Final$CTRL_GT_RND2_MEAN[i] <- NA
  } else{
    Analysis_Final$CTRL_GT_RND2_MEAN[i] <- as.numeric(mean(x2))
  }
  if(sum(is.na(c(Analysis_Final$CTRL_GT_RND1_MEAN[i], Analysis_Final$CTRL_GT_RND2_MEAN[i])))==0){
    Analysis_Final$CTRL_GT_RND1_2_MEAN[i] <- as.numeric(mean(c(Analysis_Final$CTRL_GT_RND1_MEAN[i], Analysis_Final$CTRL_GT_RND2_MEAN[i])))
    Analysis_Final[i, 87] <- as.numeric(sd(c(Analysis_Final$CTRL_GT_RND1_MEAN[i], Analysis_Final$CTRL_GT_RND2_MEAN[i])))
  } else{
    Analysis_Final$CTRL_GT_RND1_2_MEAN[i] <- NA
    Analysis_Final[i, 87] <- NA
  }
}
colnames(Analysis_Final)[87] <- "CTRL_GT_MEAN_RND1_2_SD"
```

* The mean and SD of Normalized generation time (**LSC GT mean**) at **150mM acetic acid** re-calculation

```{r}
for(i in 1:nrow(Analysis_Final)){
  x1 <- as.numeric(Analysis_Final[i, 22:24][which(!is.na(Analysis_Final[i, 22:24]))])
  x2 <- as.numeric(Analysis_Final[i, 25:27][which(!is.na(Analysis_Final[i, 25:27]))])
  if(length(x1)==0){
    Analysis_Final$AA_GT_RND1_MEAN[i] <- NA
  } else{
    Analysis_Final$AA_GT_RND1_MEAN[i] <- as.numeric(mean(x1))
  }
  if(length(x2)==0){
    Analysis_Final$AA_GT_RND2_MEAN[i] <- NA
  } else{
    Analysis_Final$AA_GT_RND2_MEAN[i] <- as.numeric(mean(x2))
  }
  if(sum(is.na(c(Analysis_Final$AA_GT_RND1_MEAN[i], Analysis_Final$AA_GT_RND2_MEAN[i])))==0){
    Analysis_Final$AA_GT_RND1_2_MEAN[i] <- as.numeric(mean(c(Analysis_Final$AA_GT_RND1_MEAN[i], Analysis_Final$AA_GT_RND2_MEAN[i])))
    Analysis_Final[i, 88] <- as.numeric(sd(c(Analysis_Final$AA_GT_RND1_MEAN[i], Analysis_Final$AA_GT_RND2_MEAN[i])))
  } else{
    Analysis_Final$AA_GT_RND1_2_MEAN[i] <- NA
    Analysis_Final[i, 88] <- NA
  }
}
colnames(Analysis_Final)[88] <- "AA_GT_MEAN_RND1_2_SD"
```

* The mean and SD of **RELATIVE** generation time (**LPI GT mean**) at **150mM acetic acid** re-calculation

```{r}
for(i in 1:nrow(Analysis_Final)){
  x1 <- as.numeric(Analysis_Final[i, 35:37][which(!is.na(Analysis_Final[i, 35:37]))])
  x2 <- as.numeric(Analysis_Final[i, 38:40][which(!is.na(Analysis_Final[i, 38:40]))])
  if(length(x1)==0){
    Analysis_Final$LPI_GT_RND1_MEAN[i] <- NA
  } else{
    Analysis_Final$LPI_GT_RND1_MEAN[i] <- as.numeric(mean(x1))
  }
  if(length(x2)==0){
    Analysis_Final$LPI_GT_RND2_MEAN[i] <- NA
  } else{
    Analysis_Final$LPI_GT_RND2_MEAN[i] <- as.numeric(mean(x2))
  }
  if(sum(is.na(c(Analysis_Final$LPI_GT_RND1_MEAN[i], Analysis_Final$LPI_GT_RND2_MEAN[i])))==0){
    Analysis_Final$LPI_GT_RND1_2_MEAN[i] <- as.numeric(mean(c(Analysis_Final$LPI_GT_RND1_MEAN[i], Analysis_Final$LPI_GT_RND2_MEAN[i])))
    Analysis_Final[i, 89] <- as.numeric(sd(c(Analysis_Final$LPI_GT_RND1_MEAN[i], Analysis_Final$LPI_GT_RND2_MEAN[i])))
  } else{
    Analysis_Final$LPI_GT_RND1_2_MEAN[i] <- NA
    Analysis_Final[i, 89] <- NA
  }
}
colnames(Analysis_Final)[89] <- "LPI_GT_MEAN_RND1_2_SD"
```

#### EXTRACT ALL LPI GT MEAN DATA POINTS WITHIN INTER-QUARTILE-RANGE (IQR)

BOX PLOT - MEAN RELATIVE GENERATION TIME (LPI GT)

```{r echo=FALSE, fig.cap="BOX PLOT - MEAN RELATIVE GENERATION TIME (LPI GT)", fig.width=2, fig.height=4}
box_stat_LPI_GT_R1_2_mean <- boxplot(Analysis_Final$LPI_GT_RND1_2_MEAN, cex=0.3)
```

Display Box-plot statistics

```{r}
box_stat_LPI_GT_R1_2_mean$stats
```

* 25th Percentile = -0.02428792
* 75th Percentile = 0.07255828

Therefore, extraction of the data points within IQR

```{r}
Intermediate_50 <- Analysis_Final$LPI_GT_RND1_2_MEAN[which(Analysis_Final$LPI_GT_RND1_2_MEAN>=-0.02428792
                                                           &Analysis_Final$LPI_GT_RND1_2_MEAN<=0.07255828)]
```

#### ESTIMATE P-VALUE

P-value is estimated by Welch two sample two-sided t-test (an adaptation of Student's t-test)

```{r}
for(i in 1:nrow(Analysis_Final)){
  if(sum(is.na(c(Analysis_Final$LPI_GT_RND1_MEAN[i], Analysis_Final$LPI_GT_RND2_MEAN[i])))==0){
    P_value <- t.test(Intermediate_50, c(Analysis_Final$LPI_GT_RND1_MEAN[i], Analysis_Final$LPI_GT_RND2_MEAN[i]))
    Analysis_Final[i, 90] <- P_value$p.value
  } else{
    Analysis_Final[i, 90] <- NA
  }
}
colnames(Analysis_Final)[90] <- "P_value_M1"
```

#### FALSE DISCOVERY RATE ADJUSTMENT OF P-VALUE 

P-value adjustment by **BENJAMINI-HOCHBERG False Discovery Rate (FDR) method**

```{r}
Analysis_Final[which(!is.na(Analysis_Final$P_value_M1)), 91] <- p.adjust(Analysis_Final$P_value_M1[which(!is.na(Analysis_Final$P_value_M1))], 
                                                                      method = "BH", 
                                                                      n = length(Analysis_Final$P_value_M1[which(!is.na(Analysis_Final$P_value_M1))]))
colnames(Analysis_Final)[91] <- "P.adjusted_M1"
```

#### P-VALUE DISGNOSTICS FOR METHOD1

NUMBER OF SIGNIFICANT STRAINS

```{r}
length(Analysis_Final$P_value_M1[which(Analysis_Final$P_value_M1<=0.05)])
length(Analysis_Final$P.adjusted_M1[which(Analysis_Final$P.adjusted_M1<=0.05)])
length(Analysis_Final$P_value_M1[which(Analysis_Final$P_value_M1<=0.1)])
length(Analysis_Final$P.adjusted_M1[which(Analysis_Final$P.adjusted_M1<=0.1)])
```

P-VALUE DIAGNOSTICS BY **HISTOGRAM ANALYSIS**

```{r echo=FALSE, fig.cap="P-VALUE DIAGNOSTICS METHOD1", fig.width=8, fig.height=8}
par(mfrow=c(2,2))
hist(Analysis_Final$P_value_M1,
     breaks = 20,
     xlab = "P-value", 
     ylab = "Frequency", 
     main = "P-value of all strains", 
     col = "skyblue",
     xlim = c(0, 1),
     ylim = c(0, 1000))
hist(Analysis_Final$P_value_M1[which(Analysis_Final$Control.gRNA==1)],
     breaks = 10,
     xlab = "P-value", 
     ylab = "Frequency", 
     main = "P-value of control strains", 
     col = "skyblue",
     xlim = c(0, 1),
     ylim = c(0, 5))
hist(Analysis_Final$P.adjusted_M1,
     breaks = 20,
     xlab = "P.adj", 
     ylab = "Frequency", 
     main = "P.adjusted values of all strains", 
     col = "skyblue",
     xlim = c(0, 1),
     ylim = c(0, 1000))
hist(Analysis_Final$P.adjusted_M1[which(Analysis_Final$Control.gRNA==1)],
     breaks = 2,
     xlab = "P.adj", 
     ylab = "Frequency", 
     main = "P.adjusted values of control strains", 
     col = "skyblue",
     xlim = c(0, 1),
     ylim = c(0, 10))
```


