---
title: "CRISPRi PHENOMICS DATA ANALYSIS"
author: "Vaskar Mukherjee"
date: "2/3/2021"
output: 
  github_document:
    toc: true
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# IMPORT SCAN-O-MATIC RAW DATA

The phenotypic data generated in scan-o-matic screening in .csv format. 
We extract both the absolute and the normalized phenotypes.

The CRISPRi strains in the library were arrayed in 24 plates in 384 format. Each CRISPRi plate was subjected to two different condition (Basal and 150 mM of Acetic acid). Therefore, for each plate four different files are generated. All files generated in a single independent experimental round are stored in a single folder. 

* **SOM_SCR_R001** : Raw data for round1

* **SOM_SCR_R002** : Raw data for round2

**ABSOLUTE DATA**

The **Absolute** dataset gives the extracted phenotypes without any spatial normalization  

**NORMALIZED DATA**

The **Normalized** dataset is generated after removal of any spatial bias. This is in log2 scale and referred as Log Strain Coefficient (LSC) values

**FILE NAMING**

Each file is named with the plate identifier in such a way so that it can be easily called programmatically  

Eg. Plate 1 absolute data in basal (Ctrl) condition have the following string  
*Ctrl1.phenotypes.Absolute*  
AND  
Plate 1 Normalized data in acetic acid (aa) stress have the string  
*aa1.phenotypes.Normalized*

## PURPOSE
At the end of this data import session, a single data.frame will be generated with the data of 24 plates. The whole dataset will be labeled with the strains attributes using the metadata key file (provided in the COMPILED_DATA folder). The data import below is shown for only Round2 dataset. Round1 can be generated modifying the folder location

**METADATA KEY FILE** : library_keyfile1536.csv

## IMPORTING THE METADATA FILE  

```{r}
Metadata_CRISPRi <- read.csv("COMPILED_DATA/library_keyfile1536.csv", na.strings = "#N/A", stringsAsFactors = FALSE)
str(Metadata_CRISPRi)
```

## GENERATE BASAL **ABSOLUTE** DATASET

```{r}
m <- vector(mode = "character", length = 0)
file.names<-vector(mode = "character", length = 0)
temp_df<-data.frame()
data_Ctrl_Abs <- data.frame()
for(i in 1:24){
  m <- paste0("Ctrl", i, ".phenotypes.Absolute") 
  file.names[i] <- dir("RAW_DATA/SOM_SCR_R002/", pattern = m, full.names = TRUE)
  temp_df <- read.csv(file.names[i], na.strings = "NoGrowth")
  data_Ctrl_Abs <- rbind(data_Ctrl_Abs, temp_df)
}
str(data_Ctrl_Abs)
```

Several phenotypes are extracted. However, the most useful for this study will be, 

* Column No: 14 i.e. **Phenotypes.ExperimentGrowthYield**
* Column No: 15 i.e. **Phenotypes.GenerationTime**

Extract only this two column in the final data.frame  
Rename the column names to prevent any ambiguity  

```{r}
data_Ctrl_Abs_Trim <- data_Ctrl_Abs[, 14:15]
colnames(data_Ctrl_Abs_Trim) <- c("CTRL_Y_ABS", "CTRL_GT_ABS")
str(data_Ctrl_Abs_Trim)
```

## GENERATE ACETIC ACID **ABSOLUTE** DATASET

Following the same strategy as above

```{r}
m <- vector(mode = "character", length = 0)
file.names<-vector(mode = "character", length = 0)
temp_df<-data.frame()
data_AA_Abs <- data.frame()
for(i in 1:24){
  m <- paste0("aa", i, ".phenotypes.Absolute") 
  file.names[i] <- dir("RAW_DATA/SOM_SCR_R002/", pattern = m, full.names = TRUE)
  temp_df <- read.csv(file.names[i], na.strings = "NoGrowth")
  data_AA_Abs <- rbind(data_AA_Abs, temp_df)
}
data_AA_Abs_Trim <- data_AA_Abs[, 14:15]
colnames(data_AA_Abs_Trim) <- c("AA_Y_ABS", "AA_GT_ABS")
```

## GENERATE BASAL **NORMALIZED** DATASET

```{r}
m <- vector(mode = "character", length = 0)
file.names<-vector(mode = "character", length = 0)
temp_df<-data.frame()
data_Ctrl_Norm <- data.frame()
for(i in 1:24){
  m <- paste0("Ctrl", i, ".phenotypes.Normalized") 
  file.names[i] <- dir("RAW_DATA/SOM_SCR_R002/", pattern = m, full.names = TRUE)
  temp_df <- read.csv(file.names[i], na.strings = "NoGrowth")
  data_Ctrl_Norm <- rbind(data_Ctrl_Norm, temp_df)
}
str(data_Ctrl_Norm)
```

The most useful for this study will be,

* Column No: 4 i.e. Phenotypes.ExperimentGrowthYield 
* Column No: 5 i.e. Phenotypes.GenerationTime

Extract only this two column  

```{r}
data_Ctrl_Norm_Trim <- data_Ctrl_Norm[, 4:5]
colnames(data_Ctrl_Norm_Trim) <- c("CTRL_Y_NORM", "CTRL_GT_NORM")
```

## GENERATE ACETIC ACID **NORMALIZED** DATASET

Same as above 

```{r}
m <- vector(mode = "character", length = 0)
file.names<-vector(mode = "character", length = 0)
temp_df<-data.frame()
data_AA_Norm <- data.frame()
for(i in 1:24){
  m <- paste0("aa", i, ".phenotypes.Normalized") 
  file.names[i] <- dir("RAW_DATA/SOM_SCR_R002/", pattern = m, full.names = TRUE)
  temp_df <- read.csv(file.names[i], na.strings = "NoGrowth")
  data_AA_Norm <- rbind(data_AA_Norm, temp_df)
}
data_AA_Norm_Trim <- data_AA_Norm[, 4:5]
colnames(data_AA_Norm_Trim) <- c("AA_Y_NORM", "AA_GT_NORM")
```

## COMBINE THE DATASETS TO OBTAIN FINAL DATAFRAME

Trimmed datasets are combined to obtain the final data.frame.
The combined data frame is labeled as data from ROUND2

```{r}
R <- rep("2nd_round", 36864)
Round_ID <- data.frame(R, stringsAsFactors = FALSE)
whole_data_R2 <- cbind(Metadata_CRISPRi, 
                       Round_ID, 
                       data_Ctrl_Abs_Trim, 
                       data_AA_Abs_Trim, 
                       data_Ctrl_Norm_Trim, 
                       data_AA_Norm_Trim)
colnames(whole_data_R2)[12] <- "Round_ID"
str(whole_data_R2)
```

## IMPORT RESULTS FROM ROUND1

The results from Round1 is already compiled to a .csv file in COMPILED_DATA folder
**Results 1st Round** : 20190903_CRISPRi_Screen_aa_1st_round.csv

Import the dataset and label as data from ROUND1

```{r}
First_round <- read.csv("COMPILED_DATA/20190903_CRISPRi_Screen_aa_1st_round.csv", 
                        na.strings = c("#N/A", "NoGrowth"), 
                        stringsAsFactors = FALSE)
R <- rep("1st_round", 36864)
Round_ID <- data.frame(R, stringsAsFactors = FALSE)
whole_data_R1 <- cbind(Metadata_CRISPRi, Round_ID, First_round[, 12:19])
colnames(whole_data_R1)[12] <- "Round_ID"
str(whole_data_R1)
```

## COMBINE THE DATASETS of ROUND 1 AND 2

```{r}
whole_data_CRISPRi_aa <- rbind(whole_data_R1, whole_data_R2)
```

# SCAN-O-MATIC PHENOMICS ANALYSIS 

In this study most of the downstream analysis was performed using the phenotype Generation_time(GT)

## PERFORM PLATE-WISE BATCH CORRECTION

Plate-wise batch correction was conducted by subtracting the median of LSC GT values (see [# IMPORT SCAN-O-MATIC RAW DATA]) of all the individual colonies on a plate from the individual LSC GT values of the colonies growing on that plate. 

i.e. if strainX is growing in Basal condition on plate Z, the corrected LSC_GT value for strainX in the Basal condition is the following;

* LSC_GT_Basal_Corrected~strainX~ = (LSC_GT_Basal~strainX~) - Median(LSC_GT Basal~PlateZ~)

```{r}
plate_ID <- as.character(unique(whole_data_CRISPRi_aa$SOURCEPLATEID))
whole_data_CRISPRi_aa_corrected <- whole_data_CRISPRi_aa
med_LogLSCctrl_RND1_GT <- vector(mode = "integer", length = 0)
med_LogLSCaa_RND1_GT <- vector(mode = "integer", length = 0)
med_LogLSCctrl_RND2_GT <- vector(mode = "integer", length = 0)
med_LogLSCaa_RND2_GT <- vector(mode = "integer", length = 0)
med_LogLSCctrl_RND1_Y <- vector(mode = "integer", length = 0)
med_LogLSCaa_RND1_Y <- vector(mode = "integer", length = 0)
med_LogLSCctrl_RND2_Y <- vector(mode = "integer", length = 0)
med_LogLSCaa_RND2_Y <- vector(mode = "integer", length = 0)
```
for(i in 1:24){
med_LogLSCctrl_RND1_GT[i] <- median(whole_data_CRISPRi_aa_corrected$CTRL_GT_NORM[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i]
                                                                                       & !is.na(whole_data_CRISPRi_aa_corrected$CTRL_GT_NORM) 
                                                                                         & whole_data_CRISPRi_aa_corrected$Round_ID=="1st_round")])

med_LogLSCaa_RND1_GT[i] <- median(whole_data_CRISPRi_aa_corrected$AA_GT_NORM[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i]
                                                                                     & !is.na(whole_data_CRISPRi_aa_corrected$AA_GT_NORM) 
                                                                                     & whole_data_CRISPRi_aa_corrected$Round_ID=="1st_round")])

med_LogLSCctrl_RND2_GT[i] <- median(whole_data_CRISPRi_aa_corrected$CTRL_GT_NORM[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i]
                                                                                         & !is.na(whole_data_CRISPRi_aa_corrected$CTRL_GT_NORM) 
                                                                                         & whole_data_CRISPRi_aa_corrected$Round_ID=="2nd_round")])

med_LogLSCaa_RND2_GT[i] <- median(whole_data_CRISPRi_aa_corrected$AA_GT_NORM[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i]
                                                                                     & !is.na(whole_data_CRISPRi_aa_corrected$AA_GT_NORM) 
                                                                                     & whole_data_CRISPRi_aa_corrected$Round_ID=="2nd_round")])

med_LogLSCctrl_RND1_Y[i] <- median(whole_data_CRISPRi_aa_corrected$CTRL_Y_NORM[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i]
                                                                                       & !is.na(whole_data_CRISPRi_aa_corrected$CTRL_Y_NORM) 
                                                                                       & whole_data_CRISPRi_aa_corrected$Round_ID=="1st_round")])

med_LogLSCaa_RND1_Y[i] <- median(whole_data_CRISPRi_aa_corrected$AA_Y_NORM[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i]
                                                                                   & !is.na(whole_data_CRISPRi_aa_corrected$AA_Y_NORM) 
                                                                                   & whole_data_CRISPRi_aa_corrected$Round_ID=="1st_round")])

med_LogLSCctrl_RND2_Y[i] <- median(whole_data_CRISPRi_aa_corrected$CTRL_Y_NORM[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i]
                                                                                       & !is.na(whole_data_CRISPRi_aa_corrected$CTRL_Y_NORM) 
                                                                                       & whole_data_CRISPRi_aa_corrected$Round_ID=="2nd_round")])

med_LogLSCaa_RND2_Y[i] <- median(whole_data_CRISPRi_aa_corrected$AA_Y_NORM[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i]
                                                                                   & !is.na(whole_data_CRISPRi_aa_corrected$AA_Y_NORM) 
                                                                                   & whole_data_CRISPRi_aa_corrected$Round_ID=="2nd_round")])
  
  whole_data_CRISPRi_aa_corrected[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i] &
                                          whole_data_CRISPRi_aa_corrected$Round_ID=="1st_round") , 23] <- whole_data_CRISPRi_aa_corrected[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i] &
                                                                                                                                                  whole_data_CRISPRi_aa_corrected$Round_ID=="1st_round"), 17] - med_LogLSCctrl_RND1_Y[i]
  whole_data_CRISPRi_aa_corrected[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i] &
                                          whole_data_CRISPRi_aa_corrected$Round_ID=="1st_round") , 24] <- whole_data_CRISPRi_aa_corrected[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i] &
                                                                                                                                                  whole_data_CRISPRi_aa_corrected$Round_ID=="1st_round"), 18] - med_LogLSCctrl_RND1_GT[i]
  whole_data_CRISPRi_aa_corrected[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i] &
                                          whole_data_CRISPRi_aa_corrected$Round_ID=="2nd_round") , 23] <- whole_data_CRISPRi_aa_corrected[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i] &
                                                                                                                                                  whole_data_CRISPRi_aa_corrected$Round_ID=="2nd_round"), 17] - med_LogLSCctrl_RND2_Y[i]
  whole_data_CRISPRi_aa_corrected[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i] &
                                          whole_data_CRISPRi_aa_corrected$Round_ID=="2nd_round") , 24] <- whole_data_CRISPRi_aa_corrected[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i] &
                                                                                                                                                  whole_data_CRISPRi_aa_corrected$Round_ID=="2nd_round"), 18] - med_LogLSCctrl_RND2_GT[i]
  whole_data_CRISPRi_aa_corrected[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i] &
                                          whole_data_CRISPRi_aa_corrected$Round_ID=="1st_round") , 25] <- whole_data_CRISPRi_aa_corrected[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i] &
                                                                                                                                                  whole_data_CRISPRi_aa_corrected$Round_ID=="1st_round"), 19] - med_LogLSCaa_RND1_Y[i]
  whole_data_CRISPRi_aa_corrected[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i] &
                                          whole_data_CRISPRi_aa_corrected$Round_ID=="1st_round") , 26] <- whole_data_CRISPRi_aa_corrected[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i] &
                                                                                                                                                  whole_data_CRISPRi_aa_corrected$Round_ID=="1st_round"), 20] - med_LogLSCaa_RND1_GT[i]
  whole_data_CRISPRi_aa_corrected[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i] &
                                          whole_data_CRISPRi_aa_corrected$Round_ID=="2nd_round") , 25] <- whole_data_CRISPRi_aa_corrected[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i] &
                                                                                                                                                  whole_data_CRISPRi_aa_corrected$Round_ID=="2nd_round"), 19] - med_LogLSCaa_RND2_Y[i]
  whole_data_CRISPRi_aa_corrected[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i] &
                                          whole_data_CRISPRi_aa_corrected$Round_ID=="2nd_round") , 26] <- whole_data_CRISPRi_aa_corrected[which(whole_data_CRISPRi_aa_corrected$SOURCEPLATEID==plate_ID[i] &
                                                                                                                                                  whole_data_CRISPRi_aa_corrected$Round_ID=="2nd_round"), 20] - med_LogLSCaa_RND2_GT[i]
}

# Determining the corrected LPI_Y 
whole_data_CRISPRi_aa_corrected[, 27] <- whole_data_CRISPRi_aa_corrected[, 25] - whole_data_CRISPRi_aa_corrected[, 23]
# Determining the corrected LPI_GT
whole_data_CRISPRi_aa_corrected[, 28] <- whole_data_CRISPRi_aa_corrected[, 26] - whole_data_CRISPRi_aa_corrected[, 24] 


## ESTIMATE THE LOG PHENOTYPIC INDEX (LPI) VALUES

LPI of strain is the difference of its normalized Generation_Time(GT) / Yield(Y) on acetic acid stress plate to the basal condition. It gives a relative estimate of how a strain performed under acetic acid stress relative to the basal condition.   i.e. LPI GT = AA_GT_NORM - CTRL_GT_NORM 


















